A generic and fast C++ optimization framework

### 2017-10-18 paper.tex
Fill out section 2; switch to minted for code listings.

\documentclass{article}
\usepackage{amsmath}
\usepackage{amsthm}
\usepackage{array}
\usepackage{booktabs}
\usepackage{graphicx}
\usepackage{wasysym}
\usepackage{tabularx}
\usepackage{threeparttable}
\usepackage{listings}

@@

\section{mlpack: a scalable C++ machine learning library}

\begin{itemize}
  \item Some quick background on what mlpack is and what its goals are.
Probably the list of four goals is good here (though maybe not in itemized
format).
  \item Discuss why we've chosen C++, and how policy-based design gives us some
really nice flexibility.  (Maybe there is not fully space for that.)
\end{itemize}

@@

\lstset{language=C++}
\begin{lstlisting}
class SparseFunction
{
  SparseFunction()
  {
    intercepts = arma::vec("20 12 15 100");
    bi = arma::vec("-4 -2 -3 -8");
  }

@@

  void Gradient(const arma::mat& p, const size_t i, arma::sp_mat& g)
  {
    g.zeros(arma::size(p));
    g[i] = 2 * p[i] + bi[i];
  }
}

\end{lstlisting}

>>>>

\documentclass{article}
\usepackage{amsmath}
\usepackage{amsthm}
\usepackage{array}
\usepackage{booktabs}
\usepackage{graphicx}
\usepackage{wasysym}
\usepackage{tabularx}
\usepackage{threeparttable}
\usepackage{minted}

@@

\section{The mlpack machine learning library}

mlpack is a C++ machine learning library with an emphasis on speed, flexibility,
and ease of use \cite{mlpack2013}; it has been continuously developed since
2007.  mlpack uses the Armadillo library \cite{sanderson2016armadillo} for
linear algebra and matrix primitives.  Both mlpack and Armadillo exploit C++
language features such as policy-based design and template metaprogramming to
provide compile-time optimizations that result in fast code \cite{pasc2017},
while retaining a simple and easy-to-use interface \cite{icopust2017}.

Many of the speedups in mlpack depend on the technique of {\it policy-based
design} \cite{Alexandrescu2001}.  With this design paradigm, mlpack
provides a number of classes with modular functionality.  As a simple example,
mlpack's mean shift clustering implementation (as of version 2.2.5) takes three
template parameters:

\begin{minted}{c++}
  template<bool UseKernel, typename KernelType, typename MatType>
  class MeanShift;
\end{minted}

Thus a user wishing to perform kernelized mean shift clustering with the
Gaussian kernel might simply use the class

\begin{minted}{c++}
  MeanShift<true, GaussianKernel, arma::mat>
\end{minted}

\noindent where {\tt arma::mat} is Armadillo's dense matrix type and {\tt
GaussianKernel} is a class that provides one simple {\tt Evaluate()} function
that returns the kernel value between its two arguments.  Since the {\tt
KernelType} argument is a template argument, any class can be used---it does not
need to be functionality supplied with mlpack.  Therefore, a user can simply
implement their own {\tt KernelType} class with the single {\tt Evaluate()}
function and use that as a template argument to the {\tt MeanShift} class, and
the compiler will generate a specialized implementation using the custom kernel.
Similar support could be accomplished by, e.g., function pointers (and this is
often done in other languages); however, templates allow us to avoid the runtime
cost of dereferencing the function pointer by generating code that directly
calls the correct function, which can be a non-negligible cost especially when
the function is called many times.

mlpack uses policy-based design throughout the library, so that any
functionality can be easily extended or modified by the user without needing to
dig into the internals of the code.  Our optimization framework is built around
this paradigm, allowing for fast prototyping of either new optimizers or new
objective functions.

@@

\begin{minted}{c++}
class SparseFunction
{
  SparseFunction()
  {
    intercepts = arma::vec("20 12 15 100");
    bi = arma::vec("-4 -2 -3 -8");
  }

@@

  void Gradient(const arma::mat& p, const size_t i, arma::sp_mat& g)
  {
    g.zeros(arma::size(p));
    g[i] = 2 * p[i] + bi[i];
  }
}
\end{minted}
